---
title: "Interactive word-network"
author: "Dereck de Mézquita"
date: "14/09/2020"
knit: (function(inputFile, encoding) { 
      rmarkdown::render(inputFile,
                        encoding=encoding, 
                        output_file=file.path(dirname(inputFile), "reports", "word-network-interactive.html")) })
output: 
  html_document: 
    fig_caption: yes
    number_sections: yes
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---

<style type="text/css">
.main-container {
  max-width: 60%;
  margin-left: auto;
  margin-right: auto;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Description

This document is a stand alone interactive visualisation of the same base network shown in the `word-network-analysis.html` report.

## Project structure

The goal of this project is to analyse chat data with my girlfriend; apply statistical methods, graph theory, and other data science techniques.

Please note that this project is presented in knitted interactive `.html` reports, you can obtain these by downloading them from this github repo at the directory `./reports` *or* you can visit them hosted on my website at: https://www.derecksnotes.com/sharing/data-science-portfolio/ds-NLP-network-inference/

This project is broken up into three big sub-projects. The project structure is as follows (hosted reports links):

- Text mining, statistical, and exploratory analysis: [derecksnotes.com: text-analysis.html](https://www.derecksnotes.com/sharing/data-science-portfolio/ds-NLP-network-inference/text-analysis.html)
- Word association network inference: [derecksnotes.com: word-network-inference.html](https://www.derecksnotes.com/sharing/data-science-portfolio/ds-NLP-network-inference/word-network-inference.html)
- Word association/network topological analysis and visualisation: [derecksnotes.com: word-network-analysis.html](https://www.derecksnotes.com/sharing/data-science-portfolio/ds-NLP-network-inference/word-network-analysis.html)
    - **Bonus -** interactive visualisation of network: [derecksnotes.com: word-network-interactive.html](https://www.derecksnotes.com/sharing/data-science-portfolio/ds-NLP-network-inference/word-network-interactive.html)

# Libraries

```{r libraries, message=FALSE, warning=FALSE}
library("igraph")
library("networkD3")
```

# Load data

```{r load-data, warning=FALSE, message=FALSE}
net <- readRDS("./outputs/networks.rds")$seventy_five

# dim(net$dec)
# net <- graph_from_data_frame(net, directed = TRUE)
# net <- graph_from_data_frame(net$data[, 1:2], directed = TRUE)
```

# Interactive network

```{r interactive-network, out.width="100%", out.height="70vh", fig.align='center'}
ndf <- get.data.frame(net)

# Function parameters
	# Data,
	# Source = 1,
	# Target = 2,
	# height = NULL,
	# width = NULL,
	# linkDistance = 50,
	# charge = -30,
	# fontSize = 7,
	# fontFamily = "serif",
	# linkColour = "#666",
	# nodeColour = "#3182bd",
	# opacity = 0.6,
	# zoom = F

# Plot
simpleNetwork(ndf, charge = -30, fontSize = 15, fontFamily = "sans-serif", nodeColour = "#1ebd30", opacity = 0.65, zoom = TRUE)
```

# Session information

```{r sesssion-info}
sessionInfo()
```

